---
title: "Getting started with FOSBAAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with FOSBAAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(FOSBAAS)
```

`FOSBAAS` is the companion package for the book *Fundamentals of Sports Business
Analytics and Strategy*. It provides the example data sets used in the book and
the simulation functions that built them. This vignette gives a short tour.

## Bundled data

Every data set is available by name once the package is attached. For example,
`season_data` holds three seasons of home games and simulated ticket sales:

```{r}
str(season_data)
head(season_data[, c("date", "team", "promotion", "ticketSales")])
```

The renewal panel, `customer_renewals`, is the basis for the lead-scoring
chapter:

```{r}
head(customer_renewals)
table(customer_renewals$renewed)
```

## Simulating a season

`f_build_season()` reproduces a season from a set of seeds and demand
parameters. The opponents, schedule, promotions and per-game attendance
multipliers are all simulated:

```{r}
season <- f_build_season(seed1 = 3000, season_year = 2024, seed2 = 714,
                         num_games = 81, seed3 = 366, num_bbh = 5,
                         num_con = 3, num_oth = 5, seed4 = 309, seed5 = 25,
                         mean_sales = 29000, sd_sales = 3500)

dim(season)
head(season[, c("gameNumber", "team", "dayOfWeek", "promotion", "ticketSales")])
```

Sales are capped at the venue capacity of 45,000:

```{r}
max(season$ticketSales)
```

## Operations: a queueing model

`f_get_MMKN()` solves an M/M/k/N queue — `k` servers and a finite capacity `N`
— and returns the standard performance measures. Here are two servers with a
capacity of five customers:

```{r}
f_get_MMKN(k = 2, N = 5, ta = 10, ts = 8)
```

We can also simulate the raw concession wait times that motivate such a model:

```{r}
wait <- f_get_wait_times(seed = 755, n = 300, rate1 = .03,
                         rate2 = .06, rate3 = .15)
summary(wait$totalTime)
```

## A small utility

Many helpers are tiny building blocks. `f_build_freq_table()` turns a vector of
observations into an empirical probability distribution:

```{r}
set.seed(1)
f_build_freq_table(sample(0:5, size = 200, replace = TRUE))
```

## Where to go next

- `?FOSBAAS` lists the function families and data sets.
- Each function's help page documents its parameters and shows a runnable
  example, e.g. `?f_build_season` or `?f_create_lead_scoring_data`.
- The `create_data/` directory in the source repository shows how each bundled
  data set was assembled from these functions.
